A comps-based approach for interpreting tree-based predictions with an application to the NFL draft

### 2025-07-06 articles/arxiv/paper.tex
Fix typo
add abstract

\begin{abstract}
  The abstract goes here.
\end{abstract}

@@

Table \ref{tab:top-ten} shows the adjusted QBR predictions for the top 10 predicted quarterbacks in the 2025 draft. Cam Ward is predicted to have the highest adjusted QBR, which matches the common consensus that he was the best quarterback in this draft class and the fact he was taken first overall in the draft. Dillon Gabriel and Shedeur Sanders were projected to have the second and third highest adjusted QBRs, respectively. While Dillon Gabriel was not previously expected to be drafted as high as Shedeur Sanders, he had a very consistent college career and finished the highest out of any quarterback in Heisman voting in 2025. Jackson Dart's draft stock increased substantially after the Scouting Combine and his pro day.

>>>>

\begin{abstract}
  Random forests are a powerful yet opaque machine learning predictive model. This poses challenges for interpretability, particularly in environments where decisions can be very high-stakes, like professional sports. In this paper, we present a method for interpreting random forest predictions in a sports setting through the lens of player comps. Player comps are a common practice in player scouting, where prospects are evaluated based on their perceived similarity to past players. We can define interpretable similarity scores that quantify how much each training observation contributes to a given prediction by leveraging the connection between random forests and adaptive nearest neighbors. A player’s similarity scores can be viewed as quantifiable player comps. We apply this methodology to evaluate quarterback prospects in the 2025 National Football League Draft using ESPN’s Total Quarterback Rating (QBR) per season, which has been regressed to zero based on attempts per season as the outcome variable. We show that our approach captures meaningful structure in the data while providing interpretability, allowing us to identify comps for top prospects such as Cam Ward, Jaxson Dart, Shedeur Sanders, and Dillon Gabriel.
\end{abstract}

@@

Table \ref{tab:top-ten} shows the adjusted QBR predictions for the top 10 predicted quarterbacks in the 2025 draft. Cam Ward is predicted to have the highest adjusted QBR, which matches the common consensus that he was the best quarterback in this draft class and the fact he was taken first overall in the draft. Dillon Gabriel and Shedeur Sanders were projected to have the second and third highest adjusted QBRs, respectively. While Dillon Gabriel was not previously expected to be drafted as high as Shedeur Sanders, he had a very consistent college career and finished the highest out of any quarterback in Heisman voting in 2025. Jaxson Dart's draft stock increased substantially after the Scouting Combine and his pro day.
